Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-01 fully-abstract-comp/technical-report.tex
Fix LaTeX compilation.
added linear capabilities and split/splice instructions w/ semantics

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}

@@

%%% Source dom
\newcommand{\shareddom}[1]{\mathrm{#1}}
\newcommand{\RegName}{\shareddom{RegisterName}}
\newcommand{\Addr}{\shareddom{Addr}}
\newcommand{\Seal}{\shareddom{Seal}}
\newcommand{\Perm}{\shareddom{Perm}}
\newcommand{\Caps}{\shareddom{Cap}}
\newcommand{\SealableCaps}{\shareddom{SealableCap}}
\newcommand{\Word}{\shareddom{Word}}
\newcommand{\Mem}{\shareddom{Memory}}
\newcommand{\Reg}{\shareddom{RegisterFile}}
\newcommand{\Stk}{\shareddom{Stack}}
\newcommand{\Conf}{\shareddom{Conf}}
\newcommand{\ExecConf}{\shareddom{ExecConf}}
\newcommand{\Global}{\shareddom{Global}}
\newcommand{\MemSeg}{\shareddom{MemorySegment}}
\newcommand{\StkFrame}{\shareddom{StackFrame}}
\newcommand{\Stack}{\shareddom{Stack}}

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

\newcommand{\sstk}{\shareddom{stk}}
\newcommand{\smsstk}{\shareddom{ms_{stk}}}
\newcommand{\sstkframe}{\shareddom{frame}}
\newcommand{\sopc}{\shareddom{opc}}
\newcommand{\sastk}{\shareddom{a_{stk}}}
\newcommand{\perm}{\shareddom{perm}}
\newcommand{\gl}{\shareddom{g}}
\newcommand{\base}{\shareddom{base}}
\newcommand{\aend}{\shareddom{end}}
\newcommand{\addr}{\shareddom{a}}
\newcommand{\scap}{\shareddom{c}}
\newcommand{\sms}{\shareddom{ms}}

@@

@@

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    & \aendbnf & \in & \Addr \uplus \{\infty \} \\
    &  \sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
\end{array}
\]

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}

\subsection{Operational Semantics}
\subsubsection{Notes}
Genrally:
\begin{itemize}
\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\end{itemize}

@@

\begin{align*}
  \tsem{\tgeta{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{a}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\_,\_,a) \text{ and }\\
                                  \perm \not\in \opaquePerm{}
                                \end{array} \\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tgetp{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_),\_,\_,\_)\\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
                                     \begin{cases}
                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = ((\perm,\gl),\baddr, \eaddr, \aaddr) \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \baddr \leq n_3 \tand \\
                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = ((\perm,\gl),n_3,n_4,\aaddr)
                                       \end{array} \\
                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = \seal{\perm, n_3,n_4}
                                       \end{array} \\
                                       \failed & \totherwise
                                     \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsetatob{r_1}{r_2}} = & 
                                 \begin{cases}
                                   \updPcAddr{\Phi\update{\reg.r_1}{c}} &
                                   \arraycolsep=0pt
                                   \begin{array}[t]{l}
                                     \Phi(r_2) = ((\perm,\gl),\baddr,\eaddr,\aaddr) \tand \\
                                     c = ((\perm,\gl),\baddr,\eaddr,\baddr) \\
                                     \perm \neq \enter \tand \perm \not\in \opaquePerm{}
                                   \end{array}
                                   \\
                                   \failed & \totherwise
                                 \end{cases}
\end{align*}

@@

>>>>

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}

@@

%%% Source dom
\newcommand{\shareddom}[1]{\mathrm{#1}}
\newcommand{\RegName}{\shareddom{RegisterName}}
\newcommand{\Addr}{\shareddom{Addr}}
\newcommand{\Seal}{\shareddom{Seal}}
\newcommand{\Perm}{\shareddom{Perm}}
\newcommand{\Caps}{\shareddom{Cap}}
\newcommand{\SealableCaps}{\shareddom{SealableCap}}
\newcommand{\Word}{\shareddom{Word}}
\newcommand{\Mem}{\shareddom{Memory}}
\newcommand{\Reg}{\shareddom{RegisterFile}}
\newcommand{\Stk}{\shareddom{Stack}}
\newcommand{\Conf}{\shareddom{Conf}}
\newcommand{\ExecConf}{\shareddom{ExecConf}}
\newcommand{\Global}{\shareddom{Global}}
\newcommand{\Linear}{\shareddom{Linear}}
\newcommand{\MemSeg}{\shareddom{MemorySegment}}
\newcommand{\StkFrame}{\shareddom{StackFrame}}
\newcommand{\Stack}{\shareddom{Stack}}

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\linbnf}{\shareddom{l}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

\newcommand{\sstk}{\shareddom{stk}}
\newcommand{\smsstk}{\shareddom{ms_{stk}}}
\newcommand{\sstkframe}{\shareddom{frame}}
\newcommand{\sopc}{\shareddom{opc}}
\newcommand{\sastk}{\shareddom{a_{stk}}}
\newcommand{\perm}{\shareddom{perm}}
\newcommand{\gl}{\shareddom{g}}
\newcommand{\lin}{\shareddom{l}}
\newcommand{\base}{\shareddom{base}}
\newcommand{\aend}{\shareddom{end}}
\newcommand{\addr}{\shareddom{a}}
\newcommand{\scap}{\shareddom{c}}
\newcommand{\sms}{\shareddom{ms}}

@@

%%% linearity
\newcommand{\plainlinearity}[1]{\mathrm{#1}}
\newcommand{\linear}{\plainlinearity{linear}}
\newcommand{\normal}{\plainlinearity{normal}}

@@

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
\end{array}
\]

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}
\lau{We should maybe add a picture with local/global hieararchy.}
\lau{For now, I will add the linear capabilities such that a linear capability stays linear. It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely). One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}

\subsection{Operational Semantics}
\subsubsection{Notes}
Genrally:
\begin{itemize}
\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\item Linear capabilities are cleared when they move around in memory.
\end{itemize}

@@

\begin{align*}
  \tsem{\tgeta{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{a}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,a) \text{ and }\\
                                  \perm \not\in \opaquePerm{}
                                \end{array} \\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tgetp{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_,\_),\_,\_,\_)\\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
                                     \begin{cases}
                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = ((\perm,\gl,\lin),\baddr, \eaddr, \aaddr) \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \baddr \leq n_3 \tand \\
                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = ((\perm,\gl,\lin),n_3,n_4,\aaddr)
                                       \end{array} \\
                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = \seal{\perm, n_3,n_4}
                                       \end{array} \\
                                       \failed & \totherwise
                                     \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsetatob{r_1}{r_2}} = & 
                                 \begin{cases}
                                   \updPcAddr{\Phi\update{\reg.r_1}{c}} &
                                   \arraycolsep=0pt
                                   \begin{array}[t]{l}
                                     \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                     c = ((\perm,\gl,\lin),\baddr,\eaddr,\baddr) \\
                                     \perm \neq \enter \tand \perm \not\in \opaquePerm{}
                                   \end{array}
                                   \\
                                   \failed & \totherwise
                                 \end{cases}
\end{align*}

@@

$\tsplice{}{}$ we allow two capabilities to be spliced if the second one starts on the address after the first one ends. They must also agree on permission, globality and linearity (so we also allow non-linear capabilities to be spliced). We do, however, exclude enter capabilities as this may cause subtle security issues.
\begin{align*}
  \tsem{\tsplice{r_1}{r_2}} = &
                                \begin{cases}
                                  \updPcAddr{\Phi\update{r_1}{\dots}\update{r_2}{c}} &
                                  \arraycolsep=0pt
                                  \begin{array}[t]{l}
                                    \Phi(r_1) = ((\perm, \gl, \lin), \baddr_1, \eaddr_1, \aaddr_1) \\
                                    \Phi(r_2) = ((\perm, \gl, \lin), \baddr_2, \eaddr_2, \aaddr_2) \\
                                    \eaddr_1 + 1 = \baddr_2 \\
                                    \perm \neq \enter \\
                                    c = ((\perm, \gl, \lin), \baddr_1, eaddr_2, \aaddr_1)
                                  \end{array}
                                  \\
                                  \failed & \totherwise
                                \end{cases}
\end{align*}

$\tsplit{}{}$ this instruction splits a capability in two. We require the address we wish to split on to be in the interval of the authority granted by the capability to prevent handing out authority from capabilities with no authority. We prohibit enter capabilities to prevent subtle errors. 
\begin{align*}
  \tsem{\tsplit{r_1}{r_2}} = &
                               \begin{cases}
                                 \updPcAddr{\Phi\update{r_1}{c_1}\update{r_2}{c_2}} &
                                 \arraycolsep=0pt
                                 \begin{array}[t]{l}
                                   \Phi(r_1) = ((\perm, \gl, \lin), \baddr, eaddr, \aaddr) \\
                                   r_2 = n \text{ or } \Phi(r_2) = n \text{ and in either case } n \in \nats\\
                                   \perm \neq \enter\\
                                   \baddr \leq n  < \eaddr\\
                                   c_1 = ((\perm,\gl,\lin),\baddr, n, \aaddr) \\
                                   c_2 = ((\perm,\gl,\lin),n + 1, \eaddr,\aaddr)
                                 \end{array} \\
                                 \failed & \totherwise
                               \end{cases}
\end{align*}
